---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---
***  
## Loading and preprocessing the data
1. Download the file and read the CSV from the zip file.
```{r get_data,cache=TRUE}
temp <- tempfile()
download.file("http://d396qusza40orc.cloudfront.net/repdata%2Fdata%2Factivity.zip",temp)
activityData <- read.csv(unz(temp, "activity.csv"))
unlink(temp)
```

2. Review the data
```{r reviewData}
head(activityData)
```


## What is mean total number of steps taken per day?
1. Summarize the total steps per day
```{r meanSteps,message=FALSE,warning=FALSE,cache=TRUE}
library(dplyr)
totalStepsPerDay <- activityData %>%
  group_by(date) %>%
  summarize(totSteps = sum(steps, na.rm = T)) 
head(totalStepsPerDay)
```
2. Create the histogram
```{r stepsPerDayHist}
hist(totalStepsPerDay$totSteps, 
     main="Histogram of Total Steps per Day",
     ylab = "Frequency",
     xlab = "Total Steps per Day",
     col = "wheat")
```
  
3. Calculate the mean and median steps per day
  
```{r meanTotalSteps}
mean(totalStepsPerDay$totSteps)
``` 
  
```{r medainTotalSteps}
median(totalStepsPerDay$totSteps)
```  
  
## What is the average daily activity pattern?
1. Calculate the mean steps per interval  
```{r meanIntervalSteps,message=FALSE,warning=FALSE,cache=TRUE}
meanStepsPerInt <- activityData %>%
  group_by(interval) %>%
  summarize(meanSteps = mean(steps, na.rm = T)) 
head(meanStepsPerInt)
```
2. Create the line graph of mean steps per interval
```{r stepsPerIntLine}
plot(meanStepsPerInt, 
     type = "l",
     main = "Mean Steps per Interval",
     ylab = "Mean steps",
     xlab = "Time interval",
     col = "blue")
```
  
3. Calculate the interval with max steps  
```{r maxStepInt,message=FALSE,warning=FALSE}
library(dplyr)
arrange(meanStepsPerInt, desc(meanSteps))[1,]
```

## Imputing missing values
1. Calculate the number of NAs  
```{r numOfNAs}
sum(is.na(activityData$steps))
```
  
2. Strategy for imputing values: use the dataset mean for the interval 

3. Create dataset with imputed values  
```{r imputeNAs, warning=FALSE,message=FALSE}
library(sqldf)

imputedActivityData <- sqldf('select case when ad.steps != "NA" then 
                             ad.steps else mspi.meanSteps end as imputed_steps, ad.date, ad.interval 
                             from activityData ad left join meanStepsPerInt mspi 
                             on ad.interval = mspi.interval')
head(imputedActivityData)

```

4. Summarize the imputed total steps per day
```{r imputedMeanSteps,message=FALSE,warning=FALSE,cache=TRUE}
library(dplyr)
totalimpStepsPerDay <- imputedActivityData %>%
  group_by(date) %>%
  summarize(totSteps = sum(imputed_steps, na.rm = T)) 
head(totalimpStepsPerDay)
```
  
5. Create histogram  
```{r imputedStepsPerDayHist}
hist(totalimpStepsPerDay$totSteps, 
     main="Histogram of Imputed Total Steps per Day",
     ylab = "Frequency",
     xlab = "Total Imputed Steps per Day",
     col = "wheat")
```
  
6. Calculate the mean and median imputed steps per day
  
```{r meanTotalimputedSteps}
mean(totalimpStepsPerDay$totSteps)
``` 
  
```{r medainTotalimputedSteps}
median(totalimpStepsPerDay$totSteps)
```  
  
## Are there differences in activity patterns between weekdays and weekends?  
  

